---
title: "beta_mave_scores_correlations"
author: "Yves Greatti"
date: "2023-09-30"
output:
  html_document:
    css: style.css
    theme: united
    code_folding: hide
    highlight: tango
    df_print: tibble
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data Load


We are currently focusing on the BRCA1 gene and its association with the breast 
cancer phenotype.
Below, we load the variant file for the BRCA1 phenotype, sourced from Genebass:
[here](https://app.genebass.org/gene/ENSG00000012048/phenotype/categorical-BreastCancer_custom-females--custom?burdenSet=pLoF&phewasOpts=1&resultIndex=variant-manhattan&resultLayout=full)

```{r, include=FALSE}

rm(list = ls())
library(tidyverse)
library(ggcorrplot)
library(plotly)
library(kableExtra)


ouput_path <- "ouputs"
figure_path <- "figures"

source("plot_lib.R", local = knitr::knit_global())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

data <- as_tibble(read.table("../data/genbass_breastcancer_custom_BRCA1.csv",
   header = TRUE, sep = ","
))

data <- data %>%
   mutate(Protein = str_extract(hgvsp, "(?<=:).+"))

head(data) %>%
   kbl() %>%
   kable_paper("hover", full_width = F, html_font = "Cambria")
```
### Loading Mave Information


```{r echo=FALSE, message=FALSE, warning=FALSE}

mave_data <- as_tibble(read.table("../data/mave_data_brn_v2.csv",
  header = TRUE, sep = ","
))
mave_data <- mave_data %>% filter(gene == "BRCA1") %>%
  mutate(Protein = str_extract(ProteinChange, "(?<=:).+")) %>% 
  select(-1)

common_proteins <- intersect(unique(data$Protein), unique(mave_data$Protein))
data_mave_df <- data %>% inner_join(mave_data, by = "Protein")

head(mave_data) %>%
  kbl() %>%
  kable_paper("hover", full_width = F, html_font = "Cambria")


```
Total proteins analyzed:  `r length(common_proteins)`.

### Correlation Between Mave Raw Score and Beta

The association between Mave raw score and Beta for BRCA1 Breast Cancer
is weak and anti-correlated.

```{r echo=FALSE, message=FALSE, warning=FALSE}

title_txt <- "Mave Raw Score vs. Beta"
xlabel <- "Raw Score"
ylabel <- "Beta"
x_col <- "raw_score"
y_col <- "beta"
xlimits <- c(-3, 1)
xbreaks <- seq(-3, 1, by = 1)
ylimits <- c(-1, 5)
ybreaks <- seq(-1, 5, by = 1)
pt <- get_scatter_plot(data_mave_df,
  title_txt,
  xlabel,
  ylabel,
  x_corr = -1.7,
  y_corr = -1,
  alpha = 2,
  raw_score,
  beta,
  point_size = 2,
  point_color = "#00AFBB",
  font_size = 14,
  annotate_text_size = 4,
  xlimits = xlimits,
  xbreaks = xbreaks,
  ylimits = ylimits,
  ybreaks = ybreaks
)
ggplotly(pt)

```
### Correlation Between Mave Normalized Raw Score and Beta
A similar weak correlation is observed between Mave normalized raw
score and Beta for BRCA1 Breast Cancer.

```{r echo=FALSE, message=FALSE, warning=FALSE}

title_txt <- "Mave Normalized Raw Score vs. Beta"
xlabel <- "Normalized Raw Score"
ylabel <- "Beta"
x_col <- "norm_raw_score"
y_col <- "beta"
xlimits <- c(-3, 1)
xbreaks <- seq(-3, 1, by = 1)
ylimits <- c(-1, 5)
ybreaks <- seq(-1, 5, by = 1)
pt <- get_scatter_plot(data_mave_df,
  title_txt,
  xlabel,
  ylabel,
  x_corr = -1.7,
  y_corr = -1,
  alpha = 2,
  norm_raw_score,
  beta,
  point_size = 2,
  point_color = "#00AFBB",
  font_size = 14,
  annotate_text_size = 4,
  xlimits = xlimits,
  xbreaks = xbreaks,
  ylimits = ylimits,
  ybreaks = ybreaks
)
ggplotly(pt)

```

### Correlation Between Mave Fuse Score and Beta
The relationship between Mave Fuse score and Beta for BRCA1 Breast Cancer 
is weak and negatively correlated.

```{r echo=FALSE, message=FALSE, warning=FALSE}

title_txt <- "Mave Fuse Score vs. Beta"
xlabel <- "Fuse Score"
ylabel <- "Beta"
x_col <- "FUSE_score"
y_col <- "beta"
xlimits <- c(-3, 1)
xbreaks <- seq(-3, 1, by = 1)
ylimits <- c(-1, 5)
ybreaks <- seq(-1, 5, by = 1)
pt <- get_scatter_plot(data_mave_df,
  title_txt,
  xlabel,
  ylabel,
  x_corr = -1.7,
  y_corr = -1,
  alpha = 2,
  FUSE_score,
  beta,
  point_size = 2,
  point_color = "#00AFBB",
  font_size = 14,
  annotate_text_size = 4,
  xlimits = xlimits,
  xbreaks = xbreaks,
  ylimits = ylimits,
  ybreaks = ybreaks
)
ggplotly(pt)

```

### Correlation Between Mave Scaled Study FUSE Score and Beta
Similar correlation exists between Mave scaled study FUSE score and Beta
for BRCA1 for Breast Cancer.

```{r echo=FALSE, message=FALSE, warning=FALSE}

title_txt <- "Mave Scaled Fuse Score vs. Beta"
xlabel <- "Scaled Fuse Score"
ylabel <- "Beta"
x_col <- "FUSE_score_scaled_study"
y_col <- "beta"
xlimits <- c(0.2, .8)
xbreaks <- seq(0.2, 0.8, by = .2)
ylimits <- c(-1, 5)
ybreaks <- seq(-1, 5, by = 1)
pt <- get_scatter_plot(data_mave_df,
  title_txt,
  xlabel,
  ylabel,
  x_corr = 0.5,
  y_corr = 5,
  alpha = 2,
  FUSE_score_scaled_study,
  beta,
  point_size = 2,
  point_color = "#00AFBB",
  font_size = 14,
  annotate_text_size = 4,
  xlimits = xlimits,
  xbreaks = xbreaks,
  ylimits = ylimits,
  ybreaks = ybreaks
)
ggplotly(pt)

```
